The Poisson approximate likelihood compared to the particle filter

### 2024-07-22 ms.Rnw
additional naming

bake(file = "output/round_01/ovov_01_el.rds",{
  el <- measlespkg::eval_logLik(mifs_global_01, ncores=ncores, np_pf = NP_MIF, nreps=ncores)
  el
})
@

@@

bake(file = "output/round_02/ovov_02_el.rds",{
  el <- measlespkg::eval_logLik(mifs_global_02, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el
})
@

@@

bake(file = "output/round_03/ovov_03_el.rds",{
  el <- measlespkg::eval_logLik(mifs_global_03, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el
})
@

@@

\begin{table}[htbp]
  \centering
  \begin{threeparttable}
    \caption{Maximum likelihood estimation of parameters by PAL for three models \citep{wwr}, and by mif2 algorithm \citep{pomppackagepaper} for the OvOv model.
    \ed{LET'S DISCUSS WHETHER WE WANT TO SHOW JUST OV-OV HERE. IT WOULD BE SIMPLER.}
    }
    \label{tab:mlebywwr}
    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} *{3}{>{\centering\arraybackslash}p{1.5cm}} >{\centering\arraybackslash}p{2.5cm}}
      \toprule
      Parameter      & EqEq  & EqOv  & OvOv  & OvOv (\textbf{pomp})$^*$ \\
      \midrule
      \(\beta_1\)    & 12.15 & 12.74 & 11.48 & $\Sexpr{est_mle_vec["beta1[unit1]"]}$\\
      \(\beta_2\)    & 0.22  & 0.21  & 0.25  & $\Sexpr{est_mle_vec["beta2[unit1]"]}$\\
      \(\beta_3\)    & 0.34  & 0.31  & 0.35  & $\Sexpr{est_mle_vec["beta3[unit1]"]}$\\
      \(\phi\)       & 0.017 & 0.14  & 0.14  & $\Sexpr{est_mle_vec["phi[unit1]"]}$\\
      \(\rho\)       & 0.022 & 0.19  & 0.16  & $\Sexpr{est_mle_vec["rho[unit1]"]}$\\
      \(\sigma^2_q\) & n/a   & 0.042 & 0.021 & $\Sexpr{est_mle_vec["sigma_q[unit1]"]}$\\
      \(\sigma_\xi\) & n/a   & n/a   & 66.89 & $\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$\\
      \addlinespace
      \multicolumn{1}{l}{\(S_{10}\)} & \multicolumn{3}{c}{$3876549^*$} & $$\Sexpr{est_mle_vec["S10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{10}\)} & \multicolumn{3}{c}{$30351$}     & $$\Sexpr{est_mle_vec["I10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{10}\)} & \multicolumn{3}{c}{$1315221$}   & $$\Sexpr{est_mle_vec["R10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{20}\)} & \multicolumn{3}{c}{$57139612$}  & $$\Sexpr{est_mle_vec["S20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{20}\)} & \multicolumn{3}{c}{$871$}       & $$\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{20}\)} & \multicolumn{3}{c}{$302852$}    & $$\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{30}\)} & \multicolumn{3}{c}{$19573727$}  & $$\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{30}\)} & \multicolumn{3}{c}{$2550$}      & $$\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{30}\)} & \multicolumn{3}{c}{$131092$}    & $$\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$$\\
      \midrule
      \addlinespace
      \textbf{AIC}  & $98866.65$ & $15154.75$ & $13778.08$ & $\Sexpr{(-2)*((length(wwr_mle) - length(params_fixed)) - est_mle_vec["logLik"])}$\\
      \addlinespace
      \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \small
      \item[*] The initial distribution parameter \(\lambda_0 = (S_{10}, I_{10}, R_{10}, S_{20}, I_{20}, R_{20}, S_{30}, I_{30}, R_{30})\) are assumed to be \textbf{fixed} in \cite{wwr}. The results of the estimates of the OvOv (\textbf{pomp}) model are obtained by using the Iterated Filtering algorithm with 3 rounds and 100 iterations, 50,000 particles, and 36 replicates in each round with the top 12 best fits in terms of likelihood chosen to be the starting value for the next round. 
    \end{tablenotes}
  \end{threeparttable}
\end{table}

>>>>

bake(file = "output/round_01/ovov_01_el.rds",{
  el_01 <- measlespkg::eval_logLik(mifs_global_01, ncores=ncores, np_pf = NP_MIF, nreps=ncores)
  el_01
})
@

@@

bake(file = "output/round_02/ovov_02_el.rds",{
  el_02 <- measlespkg::eval_logLik(mifs_global_02, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el_02
})
@

@@

bake(file = "output/round_03/ovov_03_el.rds",{
  el_03 <- measlespkg::eval_logLik(mifs_global_03, ncores = ncores, np_pf = NP_MIF, nreps=ncores)
  el_03
})
@

@@

\begin{table}[htbp]
  \centering
  \begin{threeparttable}
    \caption{Maximum likelihood estimation of parameters by PAL for three models \citep{wwr}, and by mif2 algorithm \citep{pomppackagepaper} for the OvOv model.
    \ed{LET'S DISCUSS WHETHER WE WANT TO SHOW JUST OV-OV HERE. IT WOULD BE SIMPLER.}
    }
    \label{tab:mlebywwr}
    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} *{3}{>{\centering\arraybackslash}p{1.5cm}} >{\centering\arraybackslash}p{2.5cm}}
      \toprule
      Parameter      & EqEq  & EqOv  & OvOv  & OvOv (\textbf{pomp})$^*$ \\
      \midrule
      \(\beta_1\)    & 12.15 & 12.74 & 11.48 & $\Sexpr{est_mle_vec["beta1[unit1]"]}$\\
      \(\beta_2\)    & 0.22  & 0.21  & 0.25  & $\Sexpr{est_mle_vec["beta2[unit1]"]}$\\
      \(\beta_3\)    & 0.34  & 0.31  & 0.35  & $\Sexpr{est_mle_vec["beta3[unit1]"]}$\\
      \(\phi\)       & 0.017 & 0.14  & 0.14  & $\Sexpr{est_mle_vec["phi[unit1]"]}$\\
      \(\rho\)       & 0.022 & 0.19  & 0.16  & $\Sexpr{est_mle_vec["rho[unit1]"]}$\\
      \(\sigma^2_q\) & n/a   & 0.042 & 0.021 & $\Sexpr{est_mle_vec["sigma_q[unit1]"]}$\\
      \(\sigma_\xi\) & n/a   & n/a   & 66.89 & $\Sexpr{est_mle_vec["sigma_xi[unit1]"]}$\\
      \addlinespace
      \multicolumn{1}{l}{\(S_{10}\)} & \multicolumn{3}{c}{$3876549^*$} & $$\Sexpr{est_mle_vec["S10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{10}\)} & \multicolumn{3}{c}{$30351$}     & $$\Sexpr{est_mle_vec["I10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{10}\)} & \multicolumn{3}{c}{$1315221$}   & $$\Sexpr{est_mle_vec["R10[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{20}\)} & \multicolumn{3}{c}{$57139612$}  & $$\Sexpr{est_mle_vec["S20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{20}\)} & \multicolumn{3}{c}{$871$}       & $$\Sexpr{est_mle_vec["I20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{20}\)} & \multicolumn{3}{c}{$302852$}    & $$\Sexpr{est_mle_vec["R20[unit1]"]}$$\\
      \multicolumn{1}{l}{\(S_{30}\)} & \multicolumn{3}{c}{$19573727$}  & $$\Sexpr{est_mle_vec["S30[unit1]"]}$$\\
      \multicolumn{1}{l}{\(I_{30}\)} & \multicolumn{3}{c}{$2550$}      & $$\Sexpr{est_mle_vec["I30[unit1]"]}$$\\
      \multicolumn{1}{l}{\(R_{30}\)} & \multicolumn{3}{c}{$131092$}    & $$\Sexpr{est_mle_vec["R30[unit1]"]}$$\\
      \midrule
      \addlinespace
      \textbf{AIC}  & $98866.65$ & $15154.75$ & $13778.08$ & $\Sexpr{(-2)*((length(wwr_mle) - length(params_fixed)) - est_mle_vec["logLik"])}$\\
      \addlinespace
      \bottomrule
    \end{tabular}
    \begin{tablenotes}
      \small
      \item[*] The initial distribution parameter \(\lambda_0 = (S_{10}, I_{10}, R_{10}, S_{20}, I_{20}, R_{20}, S_{30}, I_{30}, R_{30})\) are assumed to be \textbf{fixed} in \cite{wwr}. The results of the estimates of the OvOv (\textbf{pomp}) model are obtained by using the Iterated Filtering algorithm with 3 rounds and 100 iterations, 50,000 particles, and 36 replicates in each round with the top 12 best fits in terms of likelihood chosen to be the starting value for the next round. 
    \end{tablenotes}
  \end{threeparttable}
\end{table}
